---
title: "MetaPac (12 metabolites)"
subtitle: "__assay development: m-metabolic signature__"
author: "_umahajan_"
date:  "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_notebook:
    theme: united
    number_sections: yes
    toc: yes
    toc_depth: 4
    toc_float: true
---

```{r setup, include=FALSE}
chooseCRANmirror(graphics=TRUE, ind=1)
knitr::opts_chunk$set(tidy.opts=list(width.cutoff=85),
                      tidy=TRUE, 
                      echo=TRUE, 
                      warning=FALSE, 
                      message=FALSE)
```

# load packages and datasets

```{r packages}
rm(list = ls())
##---------------------------------------------------------------
##                      required packages                      --
##---------------------------------------------------------------
scriptLibraries <-  c(
  "here",
  "tidyverse",
  "stringr",
  "ggplot2",
  "ggrepel",
  "RColorBrewer",
  "ggpubr",
  "rstatix",
  "openxlsx",
  "pheatmap",
  "ComplexHeatmap",
  "sjPlot",
  "sjmisc",
  "h2o"
)
##---------------------------------------------------------------
##                      load functions                    --
##---------------------------------------------------------------
source("~/r_functions/basicFunctions.R")
source("~/r_functions/splitViolin.R")
source("~/r_functions/ImputeTransformScale.R")
##---------------------------------------------------------------
##                        load packages                        --
##---------------------------------------------------------------
installScriptLibs(scriptLibraries)
##----------------------------------------------------------------
##                         basic themes                         --
##----------------------------------------------------------------
ggplot_theme <- theme_bw() +
  theme(
    axis.line = element_line(size = 0.75),
    axis.text = element_text(
      size = 11,
      face = "bold",
      colour = "black"
    ),
    axis.title = element_text(size = 12, face = "bold")
  )
##---------------------------------------------------------------
##                    set working directory                    --
##---------------------------------------------------------------
here::here()
# load metabolite names -----------------------------------------
metaboliteNames <- 
  read.csv("./masterTable/masterTableMetaboliteNames.csv",
           stringsAsFactors = FALSE)

## remove additional
metaboliteNames$METABOLITE_NAME <- gsub("\\ [(]additional:.*", "", 
                                        metaboliteNames$METABOLITE_NAME)

## signature metabolites ----------------------------------------
sigMetabolites <- c("X578100402",
                    "X578100603",
                    "X578100716", 
                    "X578100812")
```

## initiate h2o
```{r h2o}
##----------------------------------------------------------------
##             detect the number of cores available             --
##----------------------------------------------------------------
myCores = parallel::detectCores(all.tests = TRUE) - 1

if (myCores > 20) {
  myCores = 20
} else
  myCores = myCores


memFreeG = 50
##----------------------------------------------------------------
##                         initiate h2o                         --
##----------------------------------------------------------------
h2o.init(
  nthreads = myCores,
  min_mem_size = paste(memFreeG, "g", sep = ""),
  max_mem_size = paste(memFreeG, "g", sep = "")
)
h2o.no_progress()
h2o.removeAll()

##----------------------------------------------------------------
##                        load h2o model                        --
##----------------------------------------------------------------
mod.model <- h2o.loadModel(paste0(here(),"/h2o_results/feature_reduction/GLM_1_AutoML_20211203_140114"))

testComplete <- readRDS(paste0(here(),"/ml_dataset/ImputedTest.rds"))
test <- testComplete[!testComplete$Diagnosis %in% "Non-pancreatic control", ]

test.h2o <- as.h2o(test)

cutoff <- h2o.find_threshold_by_max_metric(h2o.performance(mod.model, test.h2o), "f1")
```

## keys
```{r}
iterations <- 100
pd <- position_dodge(0.4)
```

## Influence of spiking (Figure 5A-B)

*Statistical Significance*
1. From the (up to) 12 replicates measured from each pool, mean, SD, and variance were calculated
2. This was used to calculate the difference (fold change) between test pool and control pool for each metabolite for each interferent. Statistical significance of this fold change and the confidence interval were calculated
3. The total allowable error (TAE) is calculated according to the EP7-guidelines, separately for each metabolite and for each interferent. TAE = experimentally determined bias + 3*SD
4. From the TAE, the acceptable limit Dmax is calculated as proposed by the manufacturer. Dmax =½*TAE 
5. If the upper confidence limit of the fold change is above the Dmax, the difference is considered clinically significant
6. The EP7 guidelines do not stipulate for a potential reduction in the metabolite level (fold change < 1), however this is quite important for us: matrix suppression and other effects can reduce the metabolite levels measured
7. The clinical significance was calculated (for internal use) also employing the negative TEA => TEA = +-(bias + 3SD)
8. This results in two tolerance limits, Dmax and Dmin
9. In addition to comparing the upper confidence interval limit to Dmax, a second rule applies: if the lower confidence interval limit is lower than Dmin, the difference between test pool and control pool is considered clinically significant
10. With this calculation method, there are markedly more significant differences due to the interferents. 

```{r spiking}
## Plasma dataset
datPlasma <-
  openxlsx::read.xlsx(
    paste0(here(),"/data/Boost_20180627_Metapac_05_Spiking_Global_Profiling_Pancreas_Panel.xlsx"),
    1
  ) %>%
  drop_na(VALUE) %>%
  mutate(
    MTX_REFERENCE = sub("_[^_]+$", "", .$MTX_REFERENCE),
    POOL_TYPE = gsub(" ", ".", .$POOL_TYPE),
    SPIKING_TYPE = gsub(" ", ".", .$SPIKING_TYPE),
    VARIABLE = paste(POOL_TYPE, SPIKING_TYPE, sep = "_")
  )

## define response variable
responseVariable = "VARIABLE"

## anova loop
datSubset <- datPlasma
metabolites <- unique(datSubset$METABOLITE_NAME)
### empty dataset
summaryFCPlasma <- data.frame()
loop <- 1
## loop
for (j in metabolites) {
  datMet <- datSubset[datSubset$METABOLITE_NAME %in% j,] %>%
    select(INTERFERENT, METABOLITE_NAME, VALUE, VARIABLE) %>%
    group_by(INTERFERENT, METABOLITE_NAME, VARIABLE) %>%
    mutate(Id = row_number()) %>%
    ungroup() %>%
    spread(METABOLITE_NAME, VALUE, fill = "No") %>%
    janitor::clean_names()
  
  colnames(datMet) <- toupper(colnames(datMet))
  
  interferent <- unique(datMet$INTERFERENT)
  
  print(paste("---------", j, "---------"))
  
  for (k in interferent) {
    print(k)
    datInt <- datMet[datMet$INTERFERENT %in% k,]
    datInt[[toupper(janitor::make_clean_names(j))]] <- 
      as.numeric(datInt[[toupper(janitor::make_clean_names(j))]])
    if (length(unique(datInt$VARIABLE)) < 2) {
      next
    } else {
      anovaFormula <-
        as.formula(paste(
          toupper(janitor::make_clean_names(j)),
          "~",
          responseVariable,
          sep = ""
        ))
      
      anovaFit <- aov(anovaFormula, data = datInt)
      
      tukeyMatrix <- as.matrix(TukeyHSD(anovaFit))
      
      tukeyDF <- data.frame(tukeyMatrix[[1]])
      
      correctionFDR <-
        pairwise.t.test(datInt[[toupper(janitor::make_clean_names(j))]],
                        datInt[[responseVariable]],
                        p.adjust.method = "BH")
      
      correctionFdrDF <- data.frame(correctionFDR[["p.value"]])
      
      combinations <- rownames(tukeyMatrix[[1]])
      
      for (l in seq_along(combinations)) {
        groups <- unlist(strsplit(combinations[l], "\\-"))
        
        datFC <- datInt[datInt[[responseVariable]] %in% groups, ]
        
        datFC <-
          datFC[, c(responseVariable,
                    toupper(janitor::make_clean_names(j)))]
        
        aggFormula <- as.formula(paste(".", "~", responseVariable))
        
        countFC <- aggregate(aggFormula, datFC, mean)
        
        summaryFCPlasma[loop, 1] <-
          paste(groups[1], "=", groups[2], sep = "")
        
        summaryFCPlasma[loop, 2] <- j
        
        summaryFCPlasma[loop, 3] <-
          countFC[countFC[[responseVariable]] == groups[1], 2] /
          countFC[countFC[[responseVariable]] == groups[2], 2]
        
        summaryFCPlasma[loop, 4] <-
          tukeyDF[rownames(tukeyDF) == combinations[l], "p.adj"]
        
        summaryFCPlasma[loop, 5] <-
          correctionFdrDF[rownames(correctionFdrDF) %in% groups[1],
                          colnames(correctionFdrDF) %in% groups[2]]
        
        try(tTest <-
              t.test(datInt[datInt[[responseVariable]] %in% groups, ][[toupper(janitor::make_clean_names(j))]] ~ datInt[datInt[[responseVariable]] %in% groups, ][[responseVariable]]),
            silent = TRUE)
        
        summaryFCPlasma[loop, 6] <- tTest$statistic
        
        summaryFCPlasma[loop, 7] <- k
        
        loop <- loop + 1
      }
    }
    
  }
  
}
## format colnames
colnames(summaryFCPlasma) <-
  c("VARIABLE",
    "METABOLITE_NAME",
    "logFC",
    "P_VAL",
    "FDR",
    "T_STAT",
    "INTERFERENT")  

summaryFCPlasma$MATRIX <- "Plasma"

## subset results
rowSelector <- c("Low.pool_Test.pool=Low.pool_Control.pool",
                 "High.pool_Test.pool=High.pool_Control.pool")
subsetSummaryFCPlasma <- summaryFCPlasma[summaryFCPlasma$VARIABLE %in% rowSelector,] %>%
  mutate(POOL_TYPE = gsub("\\_.*", "", .$VARIABLE)) %>%
  # filter(FDR < 0.05 & logFC < -0.5 | FDR < 0.05 & logFC > 0.5) %>%
  select(METABOLITE_NAME, INTERFERENT, MATRIX, POOL_TYPE,logFC, FDR)


datPlasma <- datPlasma %>%
  mutate(MTX_REFERENCE = sub("_[^_]+$", "", .$MTX_REFERENCE)) %>%
  select(MTX_REFERENCE, INTERFERENT, 
         POOL_TYPE, SPIKING_TYPE, METABOLITE_NAME,
         ONTOLOGY1_NAME, ONTOLOGY2_NAME, VALUE) %>%
  group_by(MTX_REFERENCE, INTERFERENT, 
           POOL_TYPE, SPIKING_TYPE, METABOLITE_NAME,
           ONTOLOGY1_NAME, ONTOLOGY2_NAME) %>%
  add_count() %>%
  mutate(MEAN = mean(VALUE),
         SD =sd(VALUE),
         VARIENCE=var(VALUE), 
         lowerCI= Rmisc::CI(VALUE)[1],
         upperCI=Rmisc::CI(VALUE)[3]) %>%
  ungroup() %>%
  group_by(INTERFERENT, 
           POOL_TYPE, SPIKING_TYPE, METABOLITE_NAME) %>%
  mutate(bias=SimDesign::bias(MEAN, n)) %>%
  distinct(MEAN,.keep_all=TRUE) %>%
  mutate(posTAE=bias+2*SD,
         negTAE=bias-2*SD,
         Dmax=0.5*posTAE,
         Dmin=0.5*negTAE) %>%
  filter(SPIKING_TYPE != "Control.pool") %>%
  select(METABOLITE_NAME, INTERFERENT, POOL_TYPE, 
         ONTOLOGY1_NAME, ONTOLOGY2_NAME, Dmax, Dmin, lowerCI, upperCI)

datMergePlasma <- full_join(datPlasma, subsetSummaryFCPlasma,
                            by = c("METABOLITE_NAME", 
                                   "INTERFERENT",
                                   "POOL_TYPE")) %>%
  filter(FDR < 0.05 & logFC < -2 | FDR < 0.05 & logFC > 2)  %>%
  filter(Dmax > upperCI | Dmin < lowerCI)


## Saline dataset
datSaline <-
  openxlsx::read.xlsx(
    "./data/Boost_20180627_Metapac_05_Spiking_Global_Profiling_Pancreas_Panel.xlsx",
    2
  ) %>%
  drop_na(VALUE) %>%
  mutate(
    MTX_REFERENCE = sub("_[^_]+$", "", .$MTX_REFERENCE),
    POOL_TYPE = gsub(" ", ".", .$POOL_TYPE),
    SPIKING_TYPE = gsub(" ", ".", .$SPIKING_TYPE),
    VARIABLE = paste(POOL_TYPE, SPIKING_TYPE, sep = "_")
  )

## define response variable
responseVariable = "VARIABLE"

## anova loop
datSubset <- datSaline
metabolites <- unique(datSubset$METABOLITE_NAME)
### empty dataset
summaryFCSaline <- data.frame()
loop <- 1
## loop
for (j in metabolites) {
  datMet <- datSubset[datSubset$METABOLITE_NAME %in% j,] %>%
    select(INTERFERENT, METABOLITE_NAME, VALUE, VARIABLE) %>%
    group_by(INTERFERENT, METABOLITE_NAME, VARIABLE) %>%
    mutate(Id = row_number()) %>%
    ungroup() %>%
    spread(METABOLITE_NAME, VALUE, fill = "No") %>%
    janitor::clean_names()
  
  colnames(datMet) <- toupper(colnames(datMet))
  
  interferent <- unique(datMet$INTERFERENT)
  
  print(paste("---------", j, "---------"))
  
  for (k in interferent) {
    print(k)
    datInt <- datMet[datMet$INTERFERENT %in% k,]
    datInt[[toupper(janitor::make_clean_names(j))]] <- 
      as.numeric(datInt[[toupper(janitor::make_clean_names(j))]])
    if (length(unique(datInt$VARIABLE)) < 2) {
      next
    } else{
      anovaFormula <-
        as.formula(paste(
          toupper(janitor::make_clean_names(j)),
          "~",
          responseVariable,
          sep = ""
        ))
      
      anovaFit <- aov(anovaFormula, data = datInt)
      
      tukeyMatrix <- as.matrix(TukeyHSD(anovaFit))
      
      tukeyDF <- data.frame(tukeyMatrix[[1]])
      
      correctionFDR <-
        pairwise.t.test(datInt[[toupper(janitor::make_clean_names(j))]],
                        datInt[[responseVariable]],
                        p.adjust.method = "BH")
      
      correctionFdrDF <- data.frame(correctionFDR[["p.value"]])
      
      combinations <- rownames(tukeyMatrix[[1]])
      
      for (l in seq_along(combinations)) {
        groups <- unlist(strsplit(combinations[l], "\\-"))
        
        datFC <- datInt[datInt[[responseVariable]] %in% groups, ]
        
        datFC <-
          datFC[, c(responseVariable,
                    toupper(janitor::make_clean_names(j)))]
        
        aggFormula <- as.formula(paste(".", "~", responseVariable))
        
        countFC <- aggregate(aggFormula, datFC, mean)
        
        summaryFCSaline[loop, 1] <-
          paste(groups[1], "=", groups[2], sep = "")
        
        summaryFCSaline[loop, 2] <- j
        
        summaryFCSaline[loop, 3] <-
          countFC[countFC[[responseVariable]] == groups[1], 2] /
          countFC[countFC[[responseVariable]] == groups[2], 2]
        
        summaryFCSaline[loop, 4] <-
          tukeyDF[rownames(tukeyDF) == combinations[l], "p.adj"]
        
        summaryFCSaline[loop, 5] <-
          correctionFdrDF[rownames(correctionFdrDF) %in% groups[1],
                          colnames(correctionFdrDF) %in% groups[2]]
        
        try(tTest <-
              t.test(datInt[datInt[[responseVariable]] %in% groups, ][[toupper(janitor::make_clean_names(j))]] ~ datInt[datInt[[responseVariable]] %in% groups, ][[responseVariable]]),
            silent = TRUE)
        
        summaryFCSaline[loop, 6] <- tTest$statistic
        
        summaryFCSaline[loop, 7] <- k
        
        loop <- loop + 1
      }
    }
    
  }
  
}
## format colnames
colnames(summaryFCSaline) <-
  c("VARIABLE",
    "METABOLITE_NAME",
    "logFC",
    "P_VAL",
    "FDR",
    "T_STAT",
    "INTERFERENT")  

summaryFCSaline$MATRIX <- "Saline"

## subset results
rowSelector <- c("saline_Test.pool=saline_Control.pool")
subsetSummaryFCSaline <- summaryFCSaline[summaryFCSaline$VARIABLE %in% rowSelector,] %>%
  mutate(POOL_TYPE = gsub("\\_.*", "", .$VARIABLE)) %>%
  # filter(FDR < 0.05 & logFC < -0.5 | FDR < 0.05 & logFC > 0.5) %>%
  select(METABOLITE_NAME, INTERFERENT, MATRIX, POOL_TYPE,logFC, FDR)


datSaline <- datSaline %>%
  mutate(MTX_REFERENCE = sub("_[^_]+$", "", .$MTX_REFERENCE)) %>%
  select(MTX_REFERENCE, INTERFERENT, 
         POOL_TYPE, SPIKING_TYPE, METABOLITE_NAME,
         ONTOLOGY1_NAME, ONTOLOGY2_NAME, VALUE) %>%
  group_by(MTX_REFERENCE, INTERFERENT, 
           POOL_TYPE, SPIKING_TYPE, METABOLITE_NAME,
           ONTOLOGY1_NAME, ONTOLOGY2_NAME) %>%
  add_count() %>%
  mutate(MEAN = mean(VALUE),
         SD =sd(VALUE),
         VARIENCE=var(VALUE), 
         lowerCI= Rmisc::CI(VALUE)[1],
         upperCI=Rmisc::CI(VALUE)[3]) %>%
  ungroup() %>%
  group_by(INTERFERENT, 
           POOL_TYPE, SPIKING_TYPE, METABOLITE_NAME) %>%
  mutate(bias=SimDesign::bias(MEAN, n)) %>%
  distinct(MEAN,.keep_all=TRUE) %>%
  mutate(posTAE=bias+2*SD,
         negTAE=bias-2*SD,
         Dmax=0.5*posTAE,
         Dmin=0.5*negTAE) %>%
  filter(SPIKING_TYPE != "Control.pool") %>%
  select(METABOLITE_NAME, INTERFERENT, POOL_TYPE, 
         ONTOLOGY1_NAME, ONTOLOGY2_NAME, Dmax, Dmin, lowerCI, upperCI)

datMergeSaline <- full_join(datSaline, subsetSummaryFCSaline,
                            by = c("METABOLITE_NAME", 
                                   "INTERFERENT",
                                   "POOL_TYPE")) %>%
  filter(FDR < 0.05 & logFC < -2 | FDR < 0.05 & logFC > 2)  %>%
  filter(Dmax > upperCI | Dmin < lowerCI)


## plot data
datPlot <- bind_rows(datMergePlasma, datMergeSaline) %>%
  group_by(INTERFERENT, MATRIX, ONTOLOGY1_NAME) %>%
  summarize(n=n()) %>%
  mutate(INTERFERENT = str_wrap(INTERFERENT, width = 20)) 
p <- datPlot %>%
  ggplot(aes(x=INTERFERENT, y=n, fill=ONTOLOGY1_NAME)) +
  geom_bar(position="stack", stat="identity", color="black", size=0.2) +
  ggplot_theme +
  xlab("") +
  ylab("number of significant metabolites") +
  facet_wrap( ~ MATRIX, ncol = 2) +
  scale_fill_manual(values = brewer.pal(length(unique(datPlot$ONTOLOGY1_NAME)), "Set3")) +
  theme(axis.text.x = element_text(angle = 90,hjust=1,vjust=0.2))

print(p)

## save results
save_plot(
  paste0("./svg/boxplot_spiking_distribution.svg"),
  fig = p,
  width = 15,
  height = 12,
  dpi = 300
)

substractedDF <- setdiff(datMergePlasma[, colnames(datMergePlasma) %in% c("METABOLITE_NAME", "INTERFERENT")],datMergeSaline[, colnames(datMergeSaline) %in% c("METABOLITE_NAME", "INTERFERENT")])

uniqueMet <- datMergePlasma[datMergePlasma$METABOLITE_NAME %in% substractedDF$METABOLITE_NAME,]

p <- uniqueMet %>%
  group_by(INTERFERENT, POOL_TYPE, ONTOLOGY1_NAME) %>%
  summarize(n=n()) %>%
  mutate(INTERFERENT = str_wrap(INTERFERENT, width = 20)) %>%
  ggplot(aes(x=INTERFERENT, y=n, fill=ONTOLOGY1_NAME)) +
  geom_bar(position="stack", stat="identity", color="black", size=0.2) +
  ggplot_theme +
  xlab("") +
  ylab("number of significant metabolites") +
  facet_wrap( ~ POOL_TYPE, ncol = 2) +
  scale_fill_manual(values = brewer.pal(length(unique(uniqueMet$ONTOLOGY1_NAME)), "Set3")) +
  theme(axis.text.x = element_text(angle = 90,hjust=1,vjust=0.2))

print(p)

## save results
save_plot(
  paste0("./svg/boxplot_spiking_distribution_test.svg"),
  fig = p,
  width = 15,
  height = 12,
  dpi = 300
)

## metabolite signature
sigMetabolitesNames <- metaboliteNames$METABOLITE_NAME[metaboliteNames$ID %in% gsub("^X", "", sigMetabolites)]

datSigMet <- bind_rows(subsetSummaryFCPlasma, subsetSummaryFCSaline) %>%
  group_by(INTERFERENT, POOL_TYPE) %>%
  filter(METABOLITE_NAME %in% sigMetabolitesNames) %>%
  select(METABOLITE_NAME, POOL_TYPE, logFC) %>%
  mutate(METABOLITE_NAME = str_wrap(METABOLITE_NAME, width = 20),
         logFC =log2(logFC))

p <- datSigMet %>%  
  ggplot(aes(
    POOL_TYPE,
    logFC,
    group = INTERFERENT
  )) +
  geom_line(
    size=1,
    aes(group = INTERFERENT, color=INTERFERENT),
    alpha = 0.5,
    show.legend = FALSE
  ) +
  geom_point(
    size = 3,
    shape = 21,
    aes(fill = INTERFERENT),
    color = "grey45",
    show.legend = TRUE
  ) +
  scale_fill_manual(values = brewer.pal(length(unique(datSigMet$INTERFERENT)), "Set1")) +
  scale_color_manual(values = brewer.pal(length(unique(datSigMet$INTERFERENT)), "Set1")) +
  ggplot_theme +
  xlab("") +
  ylab("log fold changes compared to controls") +
  facet_wrap( ~ METABOLITE_NAME, ncol = 5) +
  theme(axis.text.x = element_text(angle = 90,hjust=1,vjust=0.2),
        strip.text.x = element_text(face = "bold")) +
  ggtitle("Influence of interferent") +
  theme(legend.position = "bottom")
## print
print(p)
## save results
save_plot(
  paste0("./svg/interferent_distribution_signature.svg"),
  fig = p,
  width = 15,
  height = 10,
  dpi = 300
)
```


## Influence of comedication: VD2
```{r load df comed}
dat <- read.table(paste0(here(),"/data/MTXH_08_PDCA_VD2_predictions.tsv"), 
                  sep= "\t", 
                  header = TRUE)
dropvariables <- colnames(dat)[!grepl("^X|CA19_9", colnames(dat))]
## colnames
colnames(dat) <- toupper(colnames(dat))
## factors
dat$DIAGNOSIS <- factor(dat$DIAGNOSIS, levels = c("Non-pancreatic control",
                                                  "Chronic pancreatitis",
                                                  "Pancreatic cancer"))
## read list of medications
medications <- read.csv("./data/medications.csv")

dat$ARZNEIMITTEL <- gsub(" ", "", dat$ARZNEIMITTEL)
dat$ARZNEIMITTEL <- gsub("[(]", "_", dat$ARZNEIMITTEL)
dat$ARZNEIMITTEL <- gsub("[)]", "", dat$ARZNEIMITTEL)

##----------------------------------------------------------------
##                  log transform metabolites                  --
##----------------------------------------------------------------
## imupte train
logDat <- ImputeTransformScale(dat, 
                               Impute= TRUE,
                               Transform = TRUE,
                               Scaling = TRUE,
                               ScaleType = "Auto",
                               drop.variables = dropvariables)
colnames(logDat) <- toupper(colnames(logDat))


## convert to h2o df
logDat.h2o <- as.h2o(logDat)

## check the performance  
pred <- h2o.predict(mod.model, logDat.h2o)
pred.summary <- as.data.frame(pred)

## merge dataset
dat <- cbind(dat, pred.summary)

## plot
for (i in colnames(medications)) {
  listMed <- unlist(medications[[i]])
  listMed <- listMed[listMed != ""]
  dat[[toupper(i)]] <- ifelse(grepl(paste(listMed, collapse ="|"), dat$ARZNEIMITTEL), "Yes", "No")
  ## plot
  p <- dat %>%
    mutate(DIAGNOSIS = str_wrap(DIAGNOSIS, width = 20)) %>%
    ggplot(aes(
      x=DIAGNOSIS,
      y=PDAC
    )) +
    geom_boxplot(aes(fill=dat[[toupper(i)]]),
                 alpha=0.5, 
                 color = "black",
                 show.legend = TRUE) +
    geom_jitter(aes(fill=dat[[toupper(i)]]), 
                position=position_jitterdodge(0.2),
                color="grey45",
                shape=21,
                size =2, 
                alpha = 0.3,
                show.legend = FALSE) +
    scale_fill_manual(values = brewer.pal(2, "Set1")[1:2]) +
    ggplot_theme +
    geom_hline(yintercept = cutoff) +
    xlab("") +
    ylab("Corrected Prediction") +
    ylim(c(0,1)) +
    stat_compare_means(aes(group=dat[[toupper(i)]]),
                       label = "..p.signif..",
                       hide.ns = TRUE) +
    ggtitle(paste0("Influence comedications:", i)) +
    EnvStats::stat_n_text() +
    labs(fill=paste0(toupper(i)))
  colname <- toupper(i)
  
  countN <- dat %>%
    group_by(DIAGNOSIS,.[[toupper(i)]]) %>%
    count(.[[toupper(i)]]) 
  
  banner(paste("count of", i))
  colnames(countN)[2] <- i
  print(countN)
  
  ## print
  print(p)
  ## save results
  save_plot(
    paste0("./svg/boxplot_", i,"_prediction.svg"),
    fig = p,
    width = 15,
    height = 12,
    dpi = 300
  )
}
```

## Influence of hemoytic, icteric and lipaemic samples (Supplementary data)
```{r load df hemo}
dat <- read.table(paste0(here(),"/data/MTXH_30_METAPAC_03_icteric_hemolytic_lipemic_samples_DATA.txt"), 
                  sep= "\t", 
                  header = TRUE)

dropvariables <- colnames(dat)[!grepl("^X|CA19_9", colnames(dat))]
## create iterations
pred.summary <- data.frame(matrix(NA, ncol=iterations, nrow=nrow(dat)))

for(i in 1:iterations){
# uniquePatients <- data.frame(SUBJECT=unique(dat$SUBJECT_ID_INVENT))
dat["CA19_9"] <- runif(nrow(dat), 2, 37)

# dat <- merge(dat, uniquePatients, by.x="SUBJECT_ID_INVENT", by.y="SUBJECT")
##----------------------------------------------------------------
##                  log transform metabolites                  --
##----------------------------------------------------------------
## imupte train
imputed_data <- ImputeTransformScale(dat, 
                               Impute= TRUE,
                               Transform = TRUE,
                               Scaling = TRUE,
                               ScaleType = "Auto",
                               drop.variables = dropvariables)
colnames(imputed_data) <- toupper(colnames(imputed_data))

## convert to h2o df
imputed_data.h2o <- as.h2o(imputed_data)
## check the performance  
pred <- h2o.predict(mod.model, imputed_data.h2o)
pred.summary.itr <- as.data.frame(pred)
pred.summary[,i] <- pred.summary.itr$PDAC
}

## summarize prediction probability
PDAC <- apply(pred.summary, 1 , mean)

##----------------------------------------------------------------
##                         plot data                            --
##----------------------------------------------------------------
logDat <- dat[, colnames(dat) %in% dropvariables]
colnames(logDat) <- toupper(colnames(logDat))
## clean vectors 
logDat$PRE_ANALYTICS <- as.factor(logDat$PRE_ANALYTICS)

## define variables
responseVariable <- "PRE_ANALYTICS"
Patient_ID <- "SAMPLE_ID"

## merge dataset
plot.dat <- cbind(logDat, PDAC)

## subset data 
columnToSelect <- c(responseVariable, Patient_ID, "PDAC")
plotHeatmap <- plot.dat[, colnames(plot.dat) %in% columnToSelect]

plotHeatmap$PRE_ANALYTICS <- factor(plotHeatmap$PRE_ANALYTICS, levels = c("normal", 
                                                                          "icteric", 
                                                                          "hemolytic", 
                                                                          "lipaemic"))

## stat
stat.test <- plotHeatmap %>%
  wilcox_test(PDAC ~ PRE_ANALYTICS) %>%
  adjust_pvalue(method="bonferroni") %>%
  add_significance() %>%
  filter(group1=="normal") %>%
  mutate(y.position=1.05)

## plot
plotHeatmap <- plotHeatmap %>%
  mutate_at(vars(matches("PRE_ANALYTICS")), as.factor)

p <- ggplot(plotHeatmap, aes(
  PRE_ANALYTICS,
  PDAC,
  color = PRE_ANALYTICS
)) +
  geom_point(
    stat = "summary",
    size = 3,
    shape = 21,
    aes(fill=PRE_ANALYTICS, group = SAMPLE_ID),
    color = "grey45",
    show.legend = FALSE,
    position = pd
  ) +
  geom_boxplot(aes(fill=PRE_ANALYTICS), 
               alpha=0.5, 
               color = "black",
               show.legend = FALSE) +
  scale_fill_manual(values = brewer.pal(length(unique(plotHeatmap$PRE_ANALYTICS)), "Set1")) +
  ggplot_theme +
  xlab("") +
  ylab("Prediction probability [class: PDAC]") +
  stat_pvalue_manual(stat.test,
                     label = "p.adj.signif", 
                     x = "group2",
                     vjust = 2,
                     size = 4,
                     show.legend = FALSE,
                     hide.ns = TRUE
  ) +
  ggtitle("Influence of hemolytic, ecteric samples") +
  geom_hline(yintercept = cutoff) +
  ylim(c(0,1)) 
## print
print(p)
## save
save_plot(
  paste0("./svg/boxplot_ecteric_prediction.svg"),
  fig = p,
  width = 12,
  height = 10,
  dpi = 300
)
```

## Influence on Fasting status (Figure 5D)
```{r load df fasting}
dat <- read.table(paste0(here(),"/data/MTXH_30_METAPAC_02_Tube_type_Fasting_Pancreas_Panel_DATA.txt"), 
                  sep= "\t", 
                  header = TRUE)
## subset dataset for tubes
tubeTypes <- c("EDTA_fasted", "EDTA_P_BF", "EDTA_P_L")
dat <- dat[dat$GROUP1 %in% tubeTypes,]

dropvariables <- colnames(dat)[!grepl("^X|CA19_9", colnames(dat))]

## create iterations
pred.summary <- data.frame(matrix(NA, ncol=iterations, nrow=nrow(dat)))

for(i in 1:iterations){
uniquePatients <- data.frame(SUBJECT=unique(dat$SUBJECT_ID_INVENT))
uniquePatients["CA19_9"] <- runif(length(uniquePatients$SUBJECT), 2, 37)

dat_to_impute <- merge(dat, uniquePatients, by.x="SUBJECT_ID_INVENT", by.y="SUBJECT")

##----------------------------------------------------------------
##                  log transform metabolites                  --
##----------------------------------------------------------------
## imupte train
imputed_data <- ImputeTransformScale(dat_to_impute, 
                               Impute= TRUE,
                               Transform = TRUE,
                               Scaling = TRUE,
                               ScaleType = "Auto",
                               drop.variables = dropvariables)
colnames(imputed_data) <- toupper(colnames(imputed_data))

## convert to h2o df
imputed_data.h2o <- as.h2o(imputed_data)
## check the performance  
pred <- h2o.predict(mod.model, imputed_data.h2o)
pred.summary.itr <- as.data.frame(pred)
pred.summary[,i] <- pred.summary.itr$PDAC
}

## summarize prediction probability
PDAC <- apply(pred.summary, 1 , mean)

##----------------------------------------------------------------
##                         plot data                            --
##----------------------------------------------------------------
logDat <- dat[, colnames(dat) %in% dropvariables]
colnames(logDat) <- toupper(colnames(logDat))
## clean vectors 
logDat$GROUP1  <- as.factor(logDat$GROUP1)

## define variables
responseVariable <- "GROUP1"
Patient_ID <- "SUBJECT_ID_INVENT"

## merge dataset
plot.dat <- cbind(logDat, PDAC)

## subset data 
columnToSelect <- c(responseVariable, Patient_ID, "PDAC")
plotHeatmap <- plot.dat[, colnames(plot.dat) %in% columnToSelect]

## stat
stat.test <- plotHeatmap %>%
  wilcox_test(PDAC ~ GROUP1) %>%
  adjust_pvalue(method="bonferroni") %>%
  add_significance() %>%
  filter(group1=="EDTA_fasted") %>%
  mutate(y.position=1.05)

## plot
plotHeatmap <- plotHeatmap %>%
  mutate_at(vars(matches("GROUP1")), as.factor)

p <- ggplot(plotHeatmap, aes(
  GROUP1,
  PDAC,
  color = GROUP1
)) +
  geom_line(
    stat = "summary",
    size = 0.5,
    aes(group = SUBJECT_ID_INVENT),
    alpha = 0.5,
    color = "grey45",
    show.legend = FALSE,
    position = pd
  ) +
  geom_point(
    stat = "summary",
    size = 3,
    shape = 21,
    aes(fill=GROUP1, group = SUBJECT_ID_INVENT),
    color = "grey45",
    show.legend = FALSE,
    position = pd
  ) +
  geom_boxplot(aes(fill=GROUP1), 
               alpha=0.5, 
               color = "black",
               show.legend = FALSE) +
  scale_fill_manual(values = brewer.pal(length(unique(plotHeatmap$GROUP1)), "Set1")) +
  ggplot_theme +
  xlab("") +
  ylab("Prediction probability [class: PDAC]") +
  stat_pvalue_manual(stat.test,
                     label = "p.adj.signif", 
                     x = "group2",
                     vjust = 2,
                     size = 4,
                     show.legend = FALSE,
                     hide.ns = FALSE
  ) +
  ggtitle("Influence of fasting status") +
  geom_hline(yintercept = cutoff) +
  ylim(c(0,1)) 
## print
print(p)
## save
save_plot(
  paste0("./svg/boxplot_fasting_prediction.svg"),
  fig = p,
  width = 12,
  height = 10,
  dpi = 300
)
```

## Influence on tube types (fig S1B)
```{r load df tube types}
dat <- read.table(paste0(here(),"/data/MTXH_30_METAPAC_02_Tube_type_Fasting_Pancreas_Panel_DATA.txt"), 
                  sep= "\t", 
                  header = TRUE)

## subset dataset for tubes
tubeTypes <- c("EDTA_fasted", "EDTA_P_BF", "EDTA_P_L")
dat <- dat[!dat$GROUP1 %in% tubeTypes,]

dropvariables <- colnames(dat)[!grepl("^X|CA19_9", colnames(dat))]

## create iterations
pred.summary <- data.frame(matrix(NA, ncol=iterations, nrow=nrow(dat)))

for(i in 1:iterations){
uniquePatients <- data.frame(SUBJECT=unique(dat$SUBJECT_ID_INVENT))
uniquePatients["CA19_9"] <- runif(length(uniquePatients$SUBJECT), 2, 37)

dat_to_impute <- merge(dat, uniquePatients, by.x="SUBJECT_ID_INVENT", by.y="SUBJECT")

##----------------------------------------------------------------
##                  log transform metabolites                  --
##----------------------------------------------------------------
## imupte train
imputed_data <- ImputeTransformScale(dat_to_impute, 
                               Impute= TRUE,
                               Transform = TRUE,
                               Scaling = TRUE,
                               ScaleType = "Auto",
                               drop.variables = dropvariables)
colnames(imputed_data) <- toupper(colnames(imputed_data))

## convert to h2o df
imputed_data.h2o <- as.h2o(imputed_data)
## check the performance  
pred <- h2o.predict(mod.model, imputed_data.h2o)
pred.summary.itr <- as.data.frame(pred)
pred.summary[,i] <- pred.summary.itr$PDAC
}

## summarize prediction probability
PDAC <- apply(pred.summary, 1 , mean)

##----------------------------------------------------------------
##                         plot data                            --
##----------------------------------------------------------------
logDat <- dat[, colnames(dat) %in% dropvariables]
colnames(logDat) <- toupper(colnames(logDat))
## clean vectors 
logDat$GROUP1  <- as.factor(logDat$GROUP1)

## define variables
responseVariable <- "GROUP1"
Patient_ID <- "SUBJECT_ID_INVENT"

## merge dataset
plot.dat <- cbind(logDat, PDAC)

## subset data 
columnToSelect <- c(responseVariable, Patient_ID, "PDAC")
plotHeatmap <- plot.dat[, colnames(plot.dat) %in% columnToSelect]

## stat
stat.test <- plotHeatmap %>%
  wilcox_test(PDAC ~ GROUP1) %>%
  adjust_pvalue(method="bonferroni") %>%
  add_significance() %>%
  filter(group1=="BD_Vac_EDTA") %>%
  mutate(y.position=1.05)

## plot
plotHeatmap <- plotHeatmap %>%
  mutate_at(vars(matches("GROUP1")), as.factor)

p <- ggplot(plotHeatmap, aes(
  GROUP1,
  PDAC,
  color = GROUP1
)) +
  geom_line(
    stat = "summary",
    size = 0.5,
    aes(group = SUBJECT_ID_INVENT),
    alpha = 0.5,
    color = "grey45",
    show.legend = FALSE,
    position = pd
  ) +
  geom_point(
    stat = "summary",
    size = 3,
    shape = 21,
    aes(fill=GROUP1, group = SUBJECT_ID_INVENT),
    color = "grey45",
    show.legend = FALSE,
    position = pd
  ) +
  geom_boxplot(aes(fill=GROUP1), 
               alpha=0.5, 
               color = "black",
               show.legend = FALSE) +
  scale_fill_manual(values = brewer.pal(length(unique(plotHeatmap$GROUP1)), "Set1")) +
  ggplot_theme +
  xlab("") +
  ylab("Prediction probability [class: PDAC]") +
  stat_pvalue_manual(stat.test,
                     label = "p.adj.signif", 
                     x = "group2",
                     vjust = 2,
                     size = 4,
                     show.legend = FALSE,
                     hide.ns = TRUE
  ) +
  ggtitle("Influence of collection tubes") +
  geom_hline(yintercept = cutoff) +
  theme(axis.text.x = element_text(angle = 90,hjust=1,vjust=0.2)) +
  ylim(c(0,1)) 
## print
print(p)
## save results
save_plot(
  paste0("./svg/boxplot_collection_tubes_prediction.svg"),
  fig = p,
  width = 15,
  height = 10,
  dpi = 300
)
```

## Influence on short term storage (Fig S2B-C)
```{r load df short}
dat <- read.table(paste0(here(),"/data/MTXH_30_METAPAC_02_short_term_storage_Pancreas_Panel_DATA.txt"), 
                  sep= "\t", 
                  header = TRUE)

dropvariables <- colnames(dat)[!grepl("^X|CA19_9", colnames(dat))]

## create iterations
pred.summary <- data.frame(matrix(ncol=iterations, nrow=nrow(dat)))

for(i in 1:iterations){
uniquePatients <- data.frame(SUBJECT=unique(dat$SUBJECT))
uniquePatients["CA19_9"] <- runif(length(uniquePatients$SUBJECT), 2, 37)

dat_to_impute <- merge(dat, uniquePatients, by="SUBJECT")

##----------------------------------------------------------------
##                  log transform metabolites                  --
##----------------------------------------------------------------
## imupte train
imputed_data <- ImputeTransformScale(dat_to_impute, 
                               Impute= TRUE,
                               Transform = TRUE,
                               Scaling = TRUE,
                               ScaleType = "Auto",
                               drop.variables = dropvariables)
colnames(imputed_data) <- toupper(colnames(imputed_data))


## convert to h2o df
imputed_data.h2o <- as.h2o(imputed_data)
## check the performance  
pred <- h2o.predict(mod.model, imputed_data.h2o)
pred.summary.itr <- as.data.frame(pred)
pred.summary[,i] <- pred.summary.itr$PDAC
}

## summarize prediction probability
PDAC <- apply(pred.summary, 1 , mean)

##----------------------------------------------------------------
##                         plot data                            --
##----------------------------------------------------------------
logDat <- dat[, colnames(dat) %in% dropvariables]
colnames(logDat) <- toupper(colnames(logDat))
## clean data 
logDat$TIME_SAMPLE_PROC <- ifelse(logDat$TIME_SAMPLE_PROC == "Control", 0, logDat$TIME_SAMPLE_PROC)

## define variables
responseVariable <- "TIME_SAMPLE_PROC"
Patient_ID <- "SUBJECT"

## merge dataset
plot.dat.complete <- cbind(logDat, PDAC)


for (plot in unique(plot.dat.complete$METAPAC_WORKPACKAGE)) {
  
  plot.dat <- plot.dat.complete[plot.dat.complete$METAPAC_WORKPACKAGE %in% plot, ]
  
  ## subset data 
  columnToSelect <- c(responseVariable, Patient_ID, "PDAC", "CONFOUNDER")
  plotHeatmap <- plot.dat[, colnames(plot.dat) %in% columnToSelect]
  
  ## clean confounder
  plotHeatmap$CONFOUNDER <- ifelse(plotHeatmap$TIME_SAMPLE_PROC != 0, gsub("\\s+[^ ]+$", "", plotHeatmap$CONFOUNDER), plotHeatmap$CONFOUNDER)
  
  plotParameters <- unique(plotHeatmap$CONFOUNDER)
  
  ## plot 
  for (i in plotParameters) {
    if (grepl("Control", i)) {next}
    if (grepl("EDTA", i)) {
      banner(i)
      grpl <- c(i, "EDTA Control")
      g_subset <- plotHeatmap[plotHeatmap$CONFOUNDER %in% grpl,]
    } else if (grepl("Heparin", i)) {
      banner(i)
      grpl <- c(i, "Heparin Control")
      g_subset <- plotHeatmap[plotHeatmap$CONFOUNDER %in% grpl,]
    } else if (grepl("Serum", i)) {
      banner(i)
      grpl <- c(i, "Serum Control")
      g_subset <- plotHeatmap[plotHeatmap$CONFOUNDER %in% grpl,]
    } 
    ## convert to numeric
    g_subset$TIME_SAMPLE_PROC <- as.numeric(g_subset$TIME_SAMPLE_PROC)
    
    ## stat 
    stat.test <- g_subset %>%
      wilcox_test(PDAC ~ TIME_SAMPLE_PROC) %>%
      adjust_pvalue(method="bonferroni") %>%
      add_significance() %>%
      add_xy_position() %>%
      filter(group1==0) %>%
      mutate(y.position=1.05) 
    ## plot
    p <- g_subset %>%
      ggplot(aes(
        factor(TIME_SAMPLE_PROC),
        PDAC
      )) +
      geom_line(
    stat = "summary",
    size = 0.5,
    aes(group = SUBJECT),
    alpha = 0.5,
    color = "grey45",
    show.legend = FALSE,
    position = pd
  ) +
      geom_point(
        stat = "summary",
        size = 3,
        shape = 21,
        aes(fill= factor(TIME_SAMPLE_PROC), group=SUBJECT),
        color = "grey45",
        show.legend = FALSE,
        position = pd
      ) +
      geom_boxplot(aes(fill= factor(TIME_SAMPLE_PROC)),
                   alpha=0.5, 
                   color = "black",
                   show.legend = FALSE) +
      scale_fill_manual(values = brewer.pal(length(unique(g_subset$TIME_SAMPLE_PROC)), "Paired")) +
      ggplot_theme +
      xlab("Time in hours") +
      ylab("Prediction probability [class: PDAC]") +
      theme(strip.text.x = element_text(face = "bold")) +
      stat_pvalue_manual(stat.test,
                         label = "p.adj.signif", 
                         x = "group2",
                         vjust = 2,
                         size = 4,
                         show.legend = FALSE,
                         hide.ns = FALSE
      ) +
      ggtitle(paste("Influence of short term storage:", i)) +
      geom_hline(yintercept = cutoff) +
  ylim(c(0,1)) 
    print(p)
    ## save results
    save_plot(
      paste0(paste0("./svg/boxplot_storage_", str_replace_all(i, " ", "_"), plot, "_prediction.svg")),
      fig = p,
      width = 8,
      height = 8,
      dpi = 300
    ) 
  }
}
```

## Influence on -20 deg storage (figure S2D)
```{r load df short -20}
dat <- read.table(paste0(here(),"/data/MTXH_30_METAPAC_01_minus_20_short_term_storage_DATA.txt"), 
                  sep= "\t", 
                  header = TRUE)

dropvariables <- colnames(dat)[!grepl("^X|CA19_9", colnames(dat))]

## create iterations
pred.summary <- data.frame(matrix(ncol=iterations, nrow=nrow(dat)))

for(i in 1:iterations){

uniquePatients <- data.frame(Subject=unique(dat$Subject))

uniquePatients["CA19_9"] <- runif(length(uniquePatients$Subject), 2, 37)

dat_to_impute <- merge(dat, uniquePatients, by="Subject")

##----------------------------------------------------------------
##                  log transform metabolites                  --
##----------------------------------------------------------------
## imupte train
imputed_data <- ImputeTransformScale(dat_to_impute, 
                               Impute= TRUE,
                               Transform = TRUE,
                               Scaling = TRUE,
                               ScaleType = "Auto",
                               drop.variables=dropvariables)
colnames(imputed_data) <- toupper(colnames(imputed_data))

## convert to h2o df
imputed_data.h2o <- as.h2o(imputed_data)
## check the performance  
pred <- h2o.predict(mod.model, imputed_data.h2o)
pred.summary.itr <- as.data.frame(pred)
pred.summary[,i] <- pred.summary.itr$PDAC
}

## summarize prediction probability
PDAC <- apply(pred.summary, 1 , mean)

##----------------------------------------------------------------
##                         plot data                            --
##----------------------------------------------------------------
logDat <- dat[, colnames(dat) %in% dropvariables]
colnames(logDat) <- toupper(colnames(logDat))
## clean vectors 
logDat$STORAGE_TIME_MINUS20 <- gsub(" weeks minus 20 degrees celcius storage", "", logDat$STORAGE_TIME_MINUS20)
logDat$STORAGE_TIME_MINUS20 <- paste0("week_", logDat$STORAGE_TIME_MINUS20)
logDat$STORAGE_TIME_MINUS20 <- as.factor(logDat$STORAGE_TIME_MINUS20)

plot.dat <- cbind(logDat, PDAC)

## define variables
responseVariable <- "STORAGE_TIME_MINUS20"
Patient_ID <- "SUBJECT"

## subset data 
columnToSelect <- c(responseVariable, Patient_ID, "PDAC")
plotHeatmap <- plot.dat[, colnames(plot.dat) %in% columnToSelect]

## stats
stat.test <- plotHeatmap %>%
  wilcox_test(PDAC ~ STORAGE_TIME_MINUS20) %>%
  adjust_pvalue(method="bonferroni") %>%
  add_significance() %>%
  filter(group1=="week_0") %>%
  mutate(y.position=1.05)

pd <- position_dodge(0.4)

## plot
p <- ggplot(plotHeatmap,
            aes(STORAGE_TIME_MINUS20,
                PDAC,
                color = STORAGE_TIME_MINUS20)) +
  geom_line(
    stat = "summary",
    size = 0.5,
    aes(group = SUBJECT),
    alpha = 0.5,
    color = "grey45",
    show.legend = FALSE,
    position = pd
  ) +
  geom_point(
    stat = "summary",
    size = 3,
    shape = 21,
    aes(fill = STORAGE_TIME_MINUS20, group = SUBJECT),
    color = "grey45",
    show.legend = FALSE,
    position = pd
  ) +
  geom_boxplot(
    aes(fill = STORAGE_TIME_MINUS20),
    alpha = 0.5,
    color = "black",
    show.legend = FALSE
  ) +
  ggplot_theme +
  xlab("") +
  ylab("Prediction probability [class: PDAC]") +
  stat_pvalue_manual(
    stat.test,
    label = "p.adj.signif",
    x = "group2",
    vjust = 2,
    size = 4,
    show.legend = FALSE,
    hide.ns = FALSE
  ) +
  ggtitle("Influence of short term storage at - 20-degC") +
  # geom_line(
  #   stat = "summary",
  #   size = 2,
  #   aes(group = 1),
  #   alpha = 1,
  #   show.legend = FALSE,
  #   color = "grey25"
  # ) +
  geom_hline(yintercept = cutoff) +
  geom_point(
    stat = "summary",
    shape = 23,
    size = 8,
    aes(group = 1, fill = STORAGE_TIME_MINUS20),
    alpha = 0.8,
    show.legend = FALSE,
    color = "grey45"
  ) +
  scale_fill_manual(values = brewer.pal(length(
    unique(plotHeatmap$STORAGE_TIME_MINUS20)
  ), "Set1"))+
  ylim(c(0,1)) 
## print
print(p)
## save results
save_plot(
  paste0("./svg/boxplot_storage_minus20_prediction.svg"),
  fig = p,
  width = 8,
  height = 8,
  dpi = 300
)
```

## Influence of other cancers (supplementary data)
```{r load df other}
dat <- read.table(paste0(here(),"/data/MTXH_30_METAPAC_04_breast_cancer_colon_cancer_DATA.txt"), 
                  sep= "\t", 
                  header = TRUE)
dropvariables <- colnames(dat)[!grepl("^X|CA19_9", colnames(dat))]
## create iterations
pred.summary <- data.frame(matrix(NA, ncol=iterations, nrow=nrow(dat)))

for(i in 1:iterations){
# uniquePatients <- data.frame(SUBJECT=unique(dat$SUBJECT_ID_INVENT))
dat["CA19_9"] <- runif(nrow(dat), 2, 37)
# dat <- merge(dat, uniquePatients, by.x="SUBJECT_ID_INVENT", by.y="SUBJECT")
##----------------------------------------------------------------
##                  log transform metabolites                  --
##----------------------------------------------------------------
## imupte train
imputed_data <- ImputeTransformScale(dat, 
                               Impute= TRUE,
                               Transform = TRUE,
                               Scaling = TRUE,
                               ScaleType = "Auto",
                               drop.variables = dropvariables)
colnames(imputed_data) <- toupper(colnames(imputed_data))

## convert to h2o df
imputed_data.h2o <- as.h2o(imputed_data)
## check the performance  
pred <- h2o.predict(mod.model, imputed_data.h2o)
pred.summary.itr <- as.data.frame(pred)
pred.summary[,i] <- pred.summary.itr$PDAC
}

## summarize prediction probability
PDAC <- apply(pred.summary, 1 , mean)

##----------------------------------------------------------------
##                         plot data                            --
##----------------------------------------------------------------
logDat <- dat[, colnames(dat) %in% dropvariables]
colnames(logDat) <- toupper(colnames(logDat))
## clean vectors 
logDat$ORGAN <- as.factor(logDat$ORGAN)

## define variables
responseVariable <- "ORGAN"
Patient_ID <- "SAMPLE_ID"

## merge dataset
plot.dat <- cbind(logDat, PDAC)

## subset data 
columnToSelect <- c(responseVariable, Patient_ID, "PDAC")
plotHeatmap <- plot.dat[, colnames(plot.dat) %in% columnToSelect]

## stat
stat.test <- plotHeatmap %>%
  wilcox_test(PDAC ~ ORGAN) %>%
  adjust_pvalue(method="bonferroni") %>%
  add_significance() %>%
  mutate(y.position=1.05)

## plot
plotHeatmap <- plotHeatmap %>%
  mutate_at(vars(matches("ORGAN")), as.factor)

p <- ggplot(plotHeatmap, aes(
  ORGAN,
  PDAC,
  color = ORGAN
)) +
  geom_point(
    stat = "summary",
    size = 3,
    shape = 21,
    aes(fill=ORGAN, group = SAMPLE_ID),
    color = "grey45",
    show.legend = FALSE,
    position = pd
  ) +
  geom_boxplot(aes(fill=ORGAN), 
               alpha=0.5, 
               color = "black",
               show.legend = FALSE) +
  scale_fill_manual(values = brewer.pal(length(unique(plotHeatmap$ORGAN)), "Set1")) +
  ggplot_theme +
  xlab("") +
  ylab("Prediction probability [class: PDAC]") +
  stat_pvalue_manual(stat.test,
                     label = "p.adj.signif", 
                     x = "group2",
                     vjust = 2,
                     size = 4,
                     show.legend = FALSE,
                     hide.ns = TRUE
  ) +
  ggtitle("Influence on other cancer moeities") +
  geom_hline(yintercept = cutoff) +
  ylim(c(0,1)) 
## print
print(p)
## save results
save_plot(
  paste0("./svg/boxplot_other_cancers_prediction.svg"),
  fig = p,
  width = 10,
  height = 10,
  dpi = 300
) 
```

<!-- ## Influence of weight loss (supplementary data) -->
<!-- ```{r load df weight loss} -->
<!-- dat <- read.table(paste0(here(),"/data/MTXH_30_METAPAC_08_Weight_loss_analysis_DATA.txt"),  -->
<!--                   sep= "\t",  -->
<!--                   header = TRUE) -->

<!-- dropvariables <- colnames(dat)[!grepl("^X|CA19_9", colnames(dat))] -->

<!-- ## create iterations -->
<!-- pred.summary <- data.frame(matrix(NA, ncol=iterations, nrow=nrow(dat))) -->

<!-- for(i in 1:iterations){ -->
<!--   dat["CA19_9"] <- runif(length(uniquePatients$SUBJECT), 2, 37) -->
<!-- ##---------------------------------------------------------------- -->
<!-- ##                  log transform metabolites                  -- -->
<!-- ##---------------------------------------------------------------- -->
<!-- ## imupte train -->
<!-- imputed_data <- ImputeTransformScale(dat,  -->
<!--                                Impute= TRUE, -->
<!--                                Transform = TRUE, -->
<!--                                Scaling = TRUE, -->
<!--                                ScaleType = "Auto", -->
<!--                                drop.variables = dropvariables) -->
<!-- colnames(imputed_data) <- toupper(colnames(imputed_data)) -->

<!-- ## convert to h2o df -->
<!-- imputed_data.h2o <- as.h2o(imputed_data) -->
<!-- ## check the performance   -->
<!-- pred <- h2o.predict(mod.model, imputed_data.h2o) -->
<!-- pred.summary.itr <- as.data.frame(pred) -->
<!-- pred.summary[,i] <- pred.summary.itr$PDAC -->
<!-- } -->

<!-- ## summarize prediction probability -->
<!-- PDAC <- apply(pred.summary, 1 , mean) -->

<!-- ##---------------------------------------------------------------- -->
<!-- ##                         plot data                            -- -->
<!-- ##---------------------------------------------------------------- -->
<!-- logDat <- dat[, colnames(dat) %in% dropvariables] -->
<!-- colnames(logDat) <- toupper(colnames(logDat)) -->
<!-- ## clean vectors  -->
<!-- #normData$PRE_ANALYTICS <- as.factor(normData$PRE_ANALYTICS) -->

<!-- responseVariable <- "WEIGHT_LOSS" -->
<!-- time <- "TIMEPOINT" -->
<!-- Patient_ID <- "OWNER_EXTERNAL_REFERENCE" -->

<!-- logDat$TIMEPOINT <- gsub("_months", "", logDat$TIMEPOINT) -->

<!-- ## merge dataset -->
<!-- plot.dat <- cbind(logDat, PDAC) -->

<!-- ## subset data  -->
<!-- columnToSelect <- c(responseVariable, time, Patient_ID, "PDAC") -->
<!-- plotHeatmap <- plot.dat[, colnames(plot.dat) %in% columnToSelect] -->

<!-- ## stat -->
<!-- # stat.test <- plotHeatmap %>% -->
<!-- #   group_by(WEIGHT_LOSS) %>% -->
<!-- #   wilcox_test(PDAC ~ TIMEPOINT) %>% -->
<!-- #   adjust_pvalue(method="bonferroni") %>% -->
<!-- #   add_significance() %>% -->
<!-- #   mutate(y.position=1.05) -->

<!-- ## plot -->
<!-- plotHeatmap <- plotHeatmap %>% -->
<!--   mutate_at(vars(matches("TIMEPOINT")), as.factor) -->

<!-- plotHeatmap$WEIGHT_LOSS <- factor(plotHeatmap$WEIGHT_LOSS, levels = c("low weight loss", "moderate weight loss", "high weight loss")) -->

<!-- p <-  -->
<!--   ggplot(plotHeatmap, -->
<!--          aes( -->
<!--            TIMEPOINT, -->
<!--            PDAC, -->
<!--            color = TIMEPOINT -->
<!--          )) + -->
<!--   geom_line( -->
<!--     stat = "summary", -->
<!--     size = 0.5, -->
<!--     aes(group = OWNER_EXTERNAL_REFERENCE), -->
<!--     alpha = 0.5, -->
<!--     color = "grey45", -->
<!--     show.legend = FALSE, -->
<!--     position = pd -->
<!--   ) + -->
<!--   geom_point( -->
<!--     stat = "summary", -->
<!--     size = 3, -->
<!--     shape = 21, -->
<!--     aes(fill=TIMEPOINT, group = OWNER_EXTERNAL_REFERENCE), -->
<!--     color = "grey45", -->
<!--     show.legend = FALSE, -->
<!--     position = pd) + -->
<!--   geom_boxplot(aes(fill=TIMEPOINT),  -->
<!--                alpha=0.5,  -->
<!--                color = "black", -->
<!--                show.legend = FALSE) + -->
<!--   scale_fill_manual(values = brewer.pal(length(unique(plotHeatmap$TIMEPOINT)), "Set1")) + -->
<!--   ggplot_theme + -->
<!--   xlab("") + -->
<!--   ylab("Prediction probability [class: PDAC]") + -->
<!--   facet_wrap( ~ WEIGHT_LOSS, ncol = 3) + -->
<!--   # stat_pvalue_manual(stat.test, -->
<!--   #                    label = "p.adj.signif",  -->
<!--   #                    x = "group2", -->
<!--   #                    vjust = 2, -->
<!--   #                    size = 4, -->
<!--   #                    show.legend = FALSE, -->
<!--   #                    hide.ns = TRUE -->
<!--   # ) + -->
<!--   ggtitle("Influence on weight loss") + -->
<!--   geom_hline(yintercept = cutoff)  -->
<!-- ## print -->
<!-- print(p) -->
<!-- ## save results -->
<!-- save_plot( -->
<!--   paste0("./svg/boxplot_weiglt_loss_prediction.svg"), -->
<!--   fig = p, -->
<!--   width = 22, -->
<!--   height = 15, -->
<!--   dpi = 300 -->
<!-- ) -->
<!-- ``` -->

# computing environment
```{r}
sessionInfo()
```

